---
title: "Assignment Task 2"
output: html_notebook
---
# 1. Object x <-1:15
Construct a 5x3 matrix for the value of x.
```dim``` function to transform an n-dimensional array by giving it a dimensions attribute. 
```{r}
a <- 1:15
dim(a) <- c(5, 3)
print(a)
```

# 2. Write a program to generate the following 4x4 matrix
```
0 0 0 0
0 0 0 0
0 0 0 0
```
```{r}
a = matrix(rep(0, 16), nrow=4 , ncol=4, byrow=TRUE)
print(a)
```

# 3.The object x<-1:8
Write a program to generate the following 3x8 matrix using "```cbind(),rbind(),rev(),rep()```"
```
1 2 3 4 5 6 7 8
8 7 6 5 4 3 2 1
3 3 3 3 3 3 3 3
```
```{r}
a <- rbind(1:8, rev(1:8), rep(3, 8))
print(a)
```

# 4. Chose two different(x,y) 4x4 matrix to perform the following arithmetic operations
a. x + y
b. x - y
c. x*y
d. x%*%y
e. Difference between (c) &(d)
f. Inverse of y
g. z<-x/y
h. Transpose of z
i. Diagonal of transpose matrix
```{r}
x <- matrix(c(9, 5, 7, 9, 2, 3, 1, 9, 9, 4, 6, 7, 5, 6, 5, 7), 4, 4, byrow=TRUE)
y <- matrix(c(5, 5, 6, 3, 9, 2, 4, 3, 2, 5, 6, 5, 16, 6, 8, 8), 4, 4, byrow=TRUE)
print(x + y)
print(x - y)
print(x * y)
print(x %*% y)
print("(c) is element by element multiplication and (d) is matrix multiplication")
print(solve(y))
z <- x / y
print(z)
print(t(z))
print(diag(t(z)))
```

# 5. Difference between (i) & (ii)
i. x<-matrix(x<-1:9,3,3,byrow = TRUE)
ii. x<-matrix(x<-1:9,3,3)
```{r}
x <- matrix(x <- 1:9, 3, 3, byrow=TRUE)
print(x)
x <- matrix(x <- 1:9, 3, 3)
print(x)
print("(i) it calculate the matrix row by row and (ii) it calculate the matrix column by column.")
```

# 6. Which function can we use for inner product and outer product in matrix calculation?
```{r}
a <- c(1, 2, 3)
b <- c(3, 2, 1)
print(a %*% b)
print(a %o% b)
print("%*% is used for inner product and %o% for outer product")
```

# 7. Explain the use of all.equal() function in Matrix with a sample program.
```all.equal()``` is utility to compare matrices testing near equality. If two matrices is identical, it return true. Otherwise, It reports the differences if they are not identical. ```identical() or ==```.
```{r}
a <- matrix(1:4, ncol=2)
b <- matrix(1:4, ncol=2)
print(all.equal(a, b))
a <- matrix(1:5, ncol=2)
b <- matrix(3:7, ncol=3)
print(all.equal(a, b))
```

# 8. Explain the use of dim() function with a sample program.
```dim()``` is used to transform an atomic vector into an n-dimensional array by giving it a dimension attribute.
```{r}
a <- 1:6
dim(a) <- c(2, 3)
print(a)
```

# 9. Use of $ operator, and square brackets [] in List with example program.
```$ or [[]] and []``` is used to subset vector inside list.
```{r}
a <- list(numbers = c(3, 5), logical = TRUE, strings = c("a", "b", "c"))
a
a[1]
a$numbers
sum(a$numbers)12
```

# 10. Explain ifelse() with example.
It is a shorthand function of the traditional if...else statement.
```{r}
a = c(2, 9, 1, 4, 8)
ifelse(a %% 2 == 0, "even", "odd")
```

# 11. Write a program to get the weight and height from the user to calculate the following cases
a. If your BMI is less than 18.5, it falls within the underweight range.
b. If your BMI is 18.5 to <25, it falls within the normal.
c. If your BMI is 25.0 to <30, it falls within the overweight range.
d. If your BMI is 30.0 or higher, it falls within the obese range.
```{r}
weight <- readline(prompt = "Enter the weight: ")
weight <- as.integer(weight)
weight
height <- readline(prompt = "Enter the height: ")
height <- as.integer(height)
height
bmi = round(weight / ((height/100)^2), 1)
bmi
ifelse(bmi >= 30.0, "Obesity", 
       ifelse(bmi < 30.0 && bmi >= 25.0, "Overweight", 
              ifelse(bmi < 25.0 && bmi >= 18.5, "Normal", "Underweight")
              )
       )
```

# 12. Difference between List and Data frame with an example program.
```list``` creates a list the same way ```c``` creates a vector. However, lists do not group together individual values; lists group together R objects, such as atomic vectors and other lists.
Data frames are the two-dimensional version of a list. It is the most useful storage structure for data analysis. It stores data in an excel spreadsheet format.
```{r}
list1 <- list(35: 53, "R", list(TRUE, FALSE))
list1
```
```{r}
df <- data.frame(face = c("three", "king", "ace"),
                 suit = c("diamonds", "hearts", "clubs"), value = c(1, 2, 3))
df
```

# 13. Difference between repeat and while loop with example program.
They are companions to the ```for``` loop. ```while``` loop will take condition to run the loop whereas ```repeat``` doesn't take condition but we need to call ```break``` to exit the loop.
```{r}
b <- 1:5
a <- 1
while(a <= 3) {
  print(b[a])
  a <- a + 1
}

c<- 5
repeat {
  print(b[c])
  c <- c - 1
  if(c <= 2) {
    break
  }
}
```

